---
title: "Tutorial: Dashboards in Shiny"
subtitle: "rstudio::conf 2017"
author: "Winston Chang and Joe Cheng"
date: "January 13, 2017"
always_allow_html: yes
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
    mathjax: null
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{css}
.countdown-timer .countdown-timer-time {
  font-family: inherit;
  font-size: 50pt;
}
.countdown-timer button {
  border: 0.25px solid #CCC;
  background-color: white;
  outline: none;
}
.countdown-timer button:active {
  border: 1px solid #CCC;
  background-color: silver;
  color: white;
}
```

## Before we begin

Make sure you get an RStudio Server Pro account and log in!

Instructions: TODO

_(Supported browsers: Chrome, Firefox, Safari, IE10+. Sorry, no MS Edge.)_

---

## Say hello

#### Please take a moment to introduce yourself to your neighbors.

- Name and affiliation
- How much experience with R
- How much experience with Shiny
- How much experience building dashboards

Over the course of the next two hours, we will be doing a bunch of coding exercises. Please help each other!

```{r echo=FALSE}
countdown::countdown_timer(180)
```

---

## What sets dashboards apart from other apps?

- Automatic updating
- Potentially many viewers looking at the same data
- May or may not be interactive
- "Ten-foot" user interface

---

## Today we're going to cover

- Server: Reactivity concepts that are general, but particularly helpful for dashboarding (Joe Cheng)
- UI: Styling your app to look like a traditional dashboard (Winston Chang)

---
class: middle

# Server

---

## Problem statement

- New data comes in, either continuously or on a schedule (whether every 15s or every 24h)
  - All active sessions should automatically receive the new data
--

- May need to be transformed, aggregated, summarized, etc.
--

- Visualize data as
  - plots
  - tables
  - scalar values (e.g. "Total unique visitors today: 45,124")
--

- Call attention to exceptional conditions (e.g. "Maximum capacity exceeded", "Monthly sales goal reached")

---

## Techniques we'll cover

- Using reactiveFileReader/reactivePoll functions to automatically pick up new data
- Creating networks of reactive expressions
- Optimizing performance

---
class: middle

## reactiveFileReader and reactivePoll

---

## User-driven events vs. data-driven events
In most Shiny apps, reactivity is driven by **user gestures** (changes to the `input` object).

In dashboards, we want to drive reactivity using **changing data sources**.

Shiny comes with two functions designed to make it easy to treat changing data as reactive sources:

- `reactiveFileReader`
- `reactivePoll`

---

## But wait...

Can't we just do this?

```{r eval=FALSE}
dataset <- reactive({
  result <- read.csv("data.csv")
{{ invalidateLater(5000)}}
  result
})

output$plot <- renderPlot({
  plot(dataset()) # or whatever
})
```

--

#### This works, but is incredibly wasteful!

Every 5 seconds we will read all the data, and replot it. If `data.csv` changes every 15 minutes on average, then 99.4% of that work will be wasted effort.

Ideally, we only want `dataset` to invalidate when the file actually changes.

---

## reactiveFileReader

Creates a reactive expression that always has the "up to date" version of your file data.

```{r eval=FALSE}
dataset <- reactiveFileReader(
  intervalMillis = 1000,
  session = session,
  filePath = "data.csv",
  readFunc = read.csv
)

output$plot <- renderPlot({
  plot(dataset()) # or whatever
})
```

---

## reactiveFileReader

Creates a reactive expression that always has the "up to date" version of your file data.

```{r eval=FALSE}
dataset <- reactiveFileReader(
{{ intervalMillis = 1000,}}
  session = session,
  filePath = "data.csv",
  readFunc = read.csv
)
```

_How often to check if the file has changed (based on last-modified time)._

---

## reactiveFileReader

Creates a reactive expression that always has the "up to date" version of your file data.

```{r eval=FALSE}
dataset <- reactiveFileReader(
  intervalMillis = 1000,
{{ session = session,}}
  filePath = "data.csv",
  readFunc = read.csv
)
```

_The user session, or `NULL`._

_If non-null, then the reactive file reader will automatically stop when its session ends._

---

## reactiveFileReader

Creates a reactive expression that always has the "up to date" version of your file data.

```{r eval=FALSE}
dataset <- reactiveFileReader(
  intervalMillis = 1000,
  session = session,
{{ filePath = "data.csv",}}
  readFunc = read.csv
)
```

_The path to the data file. Must be a single file, not a directory name or glob pattern._

---

## reactiveFileReader

Creates a reactive expression that always has the "up to date" version of your file data.

```{r eval=FALSE}
dataset <- reactiveFileReader(
  intervalMillis = 1000,
  session = session,
  filePath = "data.csv",
{{ readFunc = read.csv}}
)
```

_The function that should be called (with `filePath` as an argument) whenever the file changes. The reactive expression then assumes the function's return value._

_We're using a function from `base` here, but you can use any function, including [anonymous functions](http://adv-r.had.co.nz/Functional-programming.html#anonymous-functions)._

---

## reactiveFileReader

Creates a reactive expression that always has the "up to date" version of your file data.

```{r eval=FALSE}
dataset <- reactiveFileReader(
  intervalMillis = 1000,
  session = session,
  filePath = "data.csv",
  readFunc = read.csv,
{{ stringsAsFactors = FALSE}}
)
```

_You can also add any other named arguments, and they'll be passed to `readFunc` whenever it's invoked._

---
class: middle

## Demo: tutorial_demos/01_file_reader.R

---

## Your turn

#### Create a new app that uses reactiveFileReader against a different kind of data.

Feel free to use `tutorial_demos/01_file_reader.R` as a starting point.

Some possibilities:

- Unstructured plaintext (`readLines`)
- Excel (`readxl::read_excel`)
- RDS (`readRDS`)

```{r echo=FALSE}
countdown::countdown_timer(300)
```

---

## reactivePoll

**reactiveFileReader is limited to files on disk**. It doesn't work for non-file-based data sources like databases or APIs.

--

**reactivePoll is a generalization of reactiveFileReader.** You give it two functions.

---

## reactivePoll

Creates a reactive expression that always has the "up to date" version of your ~~file~~ data.

```{r eval=FALSE}
dataset <- reactivePoll(
  intervalMillis = 1000,
  session = session,
{{ checkFunc = ...,}}
  valueFunc = ...
)
```

_A function that can execute quickly, and merely determine if anything has changed._

---

## reactivePoll

Creates a reactive expression that always has the "up to date" version of your ~~file~~ data.

```{r eval=FALSE}
dataset <- reactivePoll(
  intervalMillis = 1000,
  session = session,
  checkFunc = ...,
{{ valueFunc = ...}}
)
```

_A function with the (potentially expensive) logic for actually reading the data._

---

## Example: Database

The "check" function might return the number of rows in a table, or the max value of a timestamp column.

The "value" function actually returns the contents of the table (or some aggregation thereof).

```{r eval=FALSE}
dataset <- reactivePoll(1000, session,
  checkFunc = function() {
    # Return the latest timestamp
    DBI::dbGetQuery(dbConn, "SELECT MAX(updated) FROM log;")
  },
  valueFunc = function() {
    # Return the 20 most recently updated rows
    DBI::dbGetQuery(dbConn,
      "SELECT * FROM log ORDER BY updated DESC LIMIT 20;"
    )
  }
)
```

---

## checkFunc details

**It is important that the `checkFunc` function actually be fast!** It will block the R process while it runs!

(The slower it is, the greater you should make the polling interval.)

--

**The `checkFunc` function should not return `TRUE` or `FALSE` for changed/unchanged.** Instead, **return a value** (like the timestamp, or the count), and reactivePoll itself will keep track of whether that value is the same as the previous value or not.

---

## Example: Web API

The `checkFunc` could do an HTTP HEAD request and check the `Last-Modified` and/or `ETag` HTTP headers.

The `valueFunc` function does a full GET.

---

## Your turn

Open `tutorial_exercises/reactive_poll.R`.

This app downloads an RSS feed and summarizes the results in a table. To refresh the results, you have to reload the page. (So 2003!)

Modify this app using `reactivePoll` so that the results are always fresh.

```{r echo=FALSE}
countdown::countdown_timer(10 * 60)
```
